Supplementary Content

### 2022-01-31 experiments/2_supplement/html_supplement.Rmd
many developments

```{r echo = FALSE}
rm(list = ls())
library(tidyverse)
library(microxanox)
library(patchwork)
library(here)
source(here("r/various_useful_functions.r"))
colfunc_CB <- colorRampPalette(c("#024F17", "#B5FFC9"))
colfunc_SB <- colorRampPalette(c("#7D1402", "#FCBEB3"))
colfunc_PB <- colorRampPalette(c("#6E0172", "#F9AEFC"))
zero <- 0
unity <- 1
```

@@

```{r repl_meth1, echo=FALSE}
repl_meth <- readRDS(here("experiments/0_ss_finding/replication_method/9_strains/data/ss_data_1e6_x2x6_factorial.RDS"))
temp_meth <- readRDS(here("experiments/0_ss_finding/temporal_method/data/9_strain_SS_data.RDS"))
## To be completed.
plot_ss_result1(repl_meth,
                result_index = 1,
                filename_prefix = NULL,
                save_image_file = FALSE)
```

## Temporal method, no diversity

@@

```{r temp_meth1, echo=FALSE}
plot_ss_result3(temp_meth, 1)
```

## Both methods overlaid, no diversity

@@

```{r both_meth1, echo=FALSE}
repl_method_1 <- repl_meth$ss_res[[1]] %>%
     mutate(time = NA,
            direction = ifelse(initial_N_CB == 1, "up", "down")) %>%
  select(names(temp_meth$ssfind_result[[1]]))

@@

@@

>>>>

```{r echo = FALSE}
rm(list = ls())
library(tidyverse)
library(microxanox)
library(patchwork)
library(here)
source(here("r/various_useful_functions.r"))
source(here("experiments/1_figures_main_ms/ms_figure_functions.r"))

#print(fig_div_vs_o2diff_1strain_7row)

colfunc_CB <- colorRampPalette(c("#024F17", "#B5FFC9"))
colfunc_SB <- colorRampPalette(c("#7D1402", "#FCBEB3"))
colfunc_PB <- colorRampPalette(c("#6E0172", "#F9AEFC"))
zero <- 0
unity <- 1
```

@@

```{r repl_meth1, echo=FALSE, eval = TRUE}
repl_meth <- readRDS(here("experiments/0_ss_finding/replication_method/9_strains/data/ss_data_1e6_x2x6_factorial.RDS"))
temp_meth <- readRDS(here("experiments/0_ss_finding/temporal_method/data/ss_data_9strains_waittime1e+06_event_definition_2.RDS"))
## To be completed.
plot_ss_result1(repl_meth,
                result_index = 1,
                filename_prefix = NULL,
                save_image_file = FALSE)
```

### Temporal method, no diversity

@@

```{r temp_meth1, echo=FALSE, eval = TRUE}
plot_ss_result3(temp_meth, 1)
```

### Both methods overlaid, no diversity

@@

```{r both_meth1, echo=FALSE, eval = TRUE}
repl_method_1 <- repl_meth$ss_res[[1]] %>%
     mutate(time = NA,
            direction = ifelse(initial_N_CB == 1, "up", "down")) %>%
  select(names(temp_meth$ssfind_result[[1]]))

@@

### Further information

Here we give some further thoughts regarding the comparison of the temporal and replication method for stable state finding.

A particularly notable difference between the two approaches is in the initial condition. In the temporal approach, the initial condition at a new level of oxygen diffusivity is the stable state at the previous level of oxygen diffusivity, while in the approach of Bush et al (2017) the initial conditions are fixed and arbitrary (though this approach is still useful for illustrating the bistability of the system). This difference in initial conditions increases the region of bistability in the temporal approach because as a tipping point approaches, the functional group that will dominate after the regime shift starts from much lower abundances than in the approach by Bush et al (2017) and thus requires stronger amelioration of the environment to take over. Nevertheless, our main conclusions about the effects of functional trait diversity are the same using either approach.

The temporal approach is particularly useful when investigating the effects of strain richness. In preliminary simulations using the Bush et al (2017) approach, we found the counterintuitive result that higher strain richness led to lower resilience. We then recalled that we chose to divide the initial total abundance of the cyanobacterial (CB) functional group equally among the strains (termed a substitutive design in biodiversity manipulation experiments). This meant that the most tolerant CB strain in a three-strain simulation had higher initial abundance than the same most tolerant CB strain in a nine-strain simulation. The greater abundance of the most CB tolerant strain made it possible for a three-strain system to develop to be oxic, whereas the nine-strain system developed to be anoxic, despite identical range of trait variation. We could have chosen to do otherwise, for example start with the least tolerant strains of cyanobacteria as more abundant when the system started with high cyanobacterial abundance, and the most tolerant strain as more abundant when the system starts with low cyanobacterial abundance. Using the temporal approach removed the need to specify initial abundances (except the very first one), and therefore results were not so determined by the assumptions used to determine the initial conditions.

@@

### Effect of shorter wait times

Wait times of 1e6 hours produce the following results (same as those presented in the main report).

```{r wait_time_1e6, eval = TRUE}
wait_time <- "1e+06"

#stab_2s <- readRDS(here(paste0("experiments/0_ss_finding/temporal_method/data/stab_data_2strains_waittime", wait_time, "_event_definition_2.RDS")))
#stab_3s <- readRDS(here(paste0("experiments/0_ss_finding/temporal_method/data/stab_data_3strains_waittime", wait_time, "_event_definition_2.RDS")))
#stab_6s <- readRDS(here(paste0("experiments/0_ss_finding/temporal_method/data/stab_data_6strains_waittime", wait_time, "_event_definition_2.RDS")))
stab_9s <- readRDS(here(paste0("experiments/0_ss_finding/temporal_method/data/stab_data_9strains_waittime", wait_time, "_event_definition_2.RDS")))

#all_stab <- stab_2s %>%
  #bind_rows(stab_2s_sub1) %>%
#  bind_rows(stab_3s) %>%
  #bind_rows(stab_3s_sub1) %>%
#  bind_rows(stab_6s) %>%
  #bind_rows(stab_6s_sub1) %>%
#  bind_rows(stab_9s)
  #bind_rows(stab_9s_sub1) %>%
  #bind_rows(stab_9s_diff3)

all_stab <- stab_9s
```

```{r wait_time_1e6_2, eval = FALSE}
p1 <- fig_div_vs_o2diff_1strain_7row(all_stab,
                                     which_strain = 9,
                                     figure_title = "1e6 wait time, 9 strains")
#p1
p2 <- fig_resilience_vs_div(all_stab, which_strain = 9, figure_title = " ") 
#p2
p1 + p2

Wait times of 1e5 produce the following results, which qualitatively and often quantitatively very similar to those for wait times of 1e6. There is a notable difference when there is variation in only SB, or CB-SB variation treatments with two or three strains:

```{r wait_time_1e5, eval = FALSE}

wait_time <- "1e+05"

stab_2s <- readRDS(here(paste0("experiments/0_ss_finding/temporal_method/data/stab_data_2strains_waittime", wait_time, "_event_definition_2.RDS")))
stab_3s <- readRDS(here(paste0("experiments/0_ss_finding/temporal_method/data/stab_data_3strains_waittime", wait_time, "_event_definition_2.RDS")))
stab_6s <- readRDS(here(paste0("experiments/0_ss_finding/temporal_method/data/stab_data_6strains_waittime", wait_time, "_event_definition_2.RDS")))
stab_9s <- readRDS(here(paste0("experiments/0_ss_finding/temporal_method/data/stab_data_9strains_waittime", wait_time, "_event_definition_2.RDS")))

all_stab <- stab_2s %>%
  #bind_rows(stab_2s_sub1) %>%
  bind_rows(stab_3s) %>%
  #bind_rows(stab_3s_sub1) %>%
  bind_rows(stab_6s) %>%
  #bind_rows(stab_6s_sub1) %>%
  bind_rows(stab_9s)
  #bind_rows(stab_9s_sub1) %>%
  #bind_rows(stab_9s_diff3)

fig_div_vs_o2diff_multistrain(all_stab = all_stab,
                  which_strains = c(2, 3, 6, 9),
                  figure_title = paste0("Wait time = ", wait_time))

Here, moderate increases in diversity have the same effect as with wait times of 1e6. However, larger increases in diversity suddenly cause a large reduction in resilience of the anoxic-oxic transition. Here is an example of the stable states when diversity is moderate:

```{r}
wait_time <- "1e+04"
ss_3s <- readRDS(here(paste0("experiments/0_ss_finding/temporal_method/data/ss_data_3strains_waittime", wait_time, "_event_definition_2.RDS")))

#max <- max(ss_3s$PB_var_gmax_s)
#f <- 0.625
#fv <- f*max

only_SB_variable <- ss_3s %>%
  filter(
    abs(CB_var_gmax_s) < 0.0001,
    abs(PB_var_gmax_s) < 0.0001
  )
only_SB_variable <- only_SB_variable[order(only_SB_variable$SB_var_gmax_s), ]

only_SB_variable$SB_var_gmax_s
```

```{r}
fig_state_vs_o2diff(only_SB_variable[13,])
```

Increases in oxygen diffusivity cause strain replacement in the SB to the more tolerant strains, which then delay the shift. When there is just a little more diversity in the SB, this happens:

```{r}
fig_state_vs_o2diff(only_SB_variable[14,])
```

We believe that this phenomenon is caused by an effect that we explain with reference to an anology of using stepping stones to cross a river. Increasing diversity in our simulations is akin to increasing the width of the river, but while keeping the number of stepping stones (strains) constant. So long as the river is not too wide, we can still cross it, but once the width of the river is such that the stepping stones are too far apart, we can't cross the river. And this happens suddenly, the width increases just a tiny bit more, and we suddenly can't cross the river when we could before.

In the simulated microbial ecosystem, the strains are the stepping stones, and it is much hard (takes longer) for strains to replace each other when there is greater distance (in trait space) between them. When they can no longer replace each other in the time available, the system behaves as if only the least tolerant, highest growth rate strain (lighted shade of colour) is present. In the case illustrated above, this leads to much lower resilience of the anoxic system state.

This phenomena is dependent on having static trait values for the strains. If strain traits were dynamic (e.g. due to evolution), then we expect this result to not occur, and rather for the effects of diversity with two or three strains to be the same as those with six or nine.

### Even shorter wait times

```{r wait_time_1e4, eval = FALSE}

wait_time <- "1e+04"

stab_2s <- readRDS(here(paste0("experiments/0_ss_finding/temporal_method/data/stab_data_2strains_waittime", wait_time, "_event_definition_2.RDS")))
stab_3s <- readRDS(here(paste0("experiments/0_ss_finding/temporal_method/data/stab_data_3strains_waittime", wait_time, "_event_definition_2.RDS")))
stab_6s <- readRDS(here(paste0("experiments/0_ss_finding/temporal_method/data/stab_data_6strains_waittime", wait_time, "_event_definition_2.RDS")))
stab_9s <- readRDS(here(paste0("experiments/0_ss_finding/temporal_method/data/stab_data_9strains_waittime", wait_time, "_event_definition_2.RDS")))

all_stab <- stab_2s %>%
  #bind_rows(stab_2s_sub1) %>%
  bind_rows(stab_3s) %>%
  #bind_rows(stab_3s_sub1) %>%
  bind_rows(stab_6s) %>%
  #bind_rows(stab_6s_sub1) %>%
  bind_rows(stab_9s)
  #bind_rows(stab_9s_sub1) %>%
  #bind_rows(stab_9s_diff3)

fig_div_vs_o2diff_multistrain(all_stab = all_stab,
                  which_strains = c(2, 3, 6, 9),
                  figure_title = paste0("Wait time = ", wait_time))

```{r wait_time_1e3, eval = FALSE}

wait_time <- "1e+03"

stab_2s <- readRDS(here(paste0("experiments/0_ss_finding/temporal_method/data/stab_data_2strains_waittime", wait_time, "_event_definition_2.RDS")))
stab_3s <- readRDS(here(paste0("experiments/0_ss_finding/temporal_method/data/stab_data_3strains_waittime", wait_time, "_event_definition_2.RDS")))
stab_6s <- readRDS(here(paste0("experiments/0_ss_finding/temporal_method/data/stab_data_6strains_waittime", wait_time, "_event_definition_2.RDS")))
stab_9s <- readRDS(here(paste0("experiments/0_ss_finding/temporal_method/data/stab_data_9strains_waittime", wait_time, "_event_definition_2.RDS")))

all_stab <- stab_2s %>%
  #bind_rows(stab_2s_sub1) %>%
  bind_rows(stab_3s) %>%
  #bind_rows(stab_3s_sub1) %>%
  bind_rows(stab_6s) %>%
  #bind_rows(stab_6s_sub1) %>%
  bind_rows(stab_9s)
  #bind_rows(stab_9s_sub1) %>%
  #bind_rows(stab_9s_diff3)

fig_div_vs_o2diff_multistrain(all_stab = all_stab,
                  which_strains = c(2, 3, 6, 9),
                  figure_title = paste0("Wait time = ", wait_time))

```{r wait_time_1e2, eval = FALSE}

wait_time <- "1e+02"

stab_2s <- readRDS(here(paste0("experiments/0_ss_finding/temporal_method/data/stab_data_2strains_waittime", wait_time, "_event_definition_2.RDS")))
stab_3s <- readRDS(here(paste0("experiments/0_ss_finding/temporal_method/data/stab_data_3strains_waittime", wait_time, "_event_definition_2.RDS")))
stab_6s <- readRDS(here(paste0("experiments/0_ss_finding/temporal_method/data/stab_data_6strains_waittime", wait_time, "_event_definition_2.RDS")))
stab_9s <- readRDS(here(paste0("experiments/0_ss_finding/temporal_method/data/stab_data_9strains_waittime", wait_time, "_event_definition_2.RDS")))

all_stab <- stab_2s %>%
  #bind_rows(stab_2s_sub1) %>%
  bind_rows(stab_3s) %>%
  #bind_rows(stab_3s_sub1) %>%
  bind_rows(stab_6s) %>%
  #bind_rows(stab_6s_sub1) %>%
  bind_rows(stab_9s)
  #bind_rows(stab_9s_sub1) %>%
  #bind_rows(stab_9s_diff3)

fig_div_vs_o2diff_multistrain(all_stab = all_stab,
                  which_strains = c(2, 3, 6, 9),
                  figure_title = paste0("Wait time = ", wait_time))

### Example transient at 1'000'000 hours
